---
title: "Homework 5"
author: "Your Name Here"
subtitle: "Winter 2025"
header-includes:
    - \usepackage{amsmath}
    - \usepackage{amsthm}
output: pdf_document
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.

knitr::opts_chunk$set(echo = TRUE, message= FALSE, warning = FALSE)
options(pillar.sigfig=6)
library(tidyverse)

```

### Instructions

- This homework is due in Gradescope on Wednesday Feb 19 by midnight PST. There is a 15 minute grace period and submissions made during this time will not be marked as late. Any work submitted past this period is considered late.  

- Please answer the following questions in the order in which they are posed. 

- Don't forget to (i) make a local copy this document for your work and to (ii) knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope.  

* * *

### Exercises

1. (MOM vs.MLE) Suppose $X_1,X_2,\dots, X_n$ are independently drawn from the PDF
$$f_{\theta}(x) = (\theta +1)x^{\theta} \ \ \ 0 \leq x \leq 1$$
where the (unknown) parameter $\theta \geq -1$ so the PDF is non-zero and integrates to 1. Say we observe $x_1,x_2,\dots,x_n$.

a. Derive $\widehat{\theta}_0^{mom}$, the method of moments estimate for $\theta_0$, the true value of the parameter.

b. Derive $\widehat{\theta}_0^{mle}$, the maximum likelihood estimate of $\theta_0$. Be sure to show
   - likelihood function
   - log-likelihood function
   - first derivative condition
   - second derivative test
   
c. Use R as a calculator to find $\widehat{\theta}_0^{mom}$ for the sample below. Please be sure to show your code and output.

    ```{r echo =T}
    x <- c(0.90, 0.78, 0.93, 0.64, 0.45, 0.85, 0.75, 0.93, 0.98, 0.78)

    ```

d. Make a plot of the log-likelihood function for the data from part c.and then use R as a calculator to find $\widehat{\theta}_0^{mle}$. Please be sure to show your code and output.


2. (Light bulbs) A set of	cheap light	bulbs have a lifetime (in hours) which is	exponentially distributed with unknown rate	$\lambda > 0$:
$$f_{\lambda}(x) = \lambda exp(-\lambda\:x), \ \  x  \geq 0$$

   Choosing	a random	sample of ten light	bulbs, they are turned on simultaneously and observed for 48 hours. During this period, six bulbs went out, at times $x_1,x_2,\dots,x_6$. At the end of the experiment, four light bulbs were still working. In other words, all we know about them is that their lifetime exceeded 48 hours.
 

a.  Formulate the likelihood function $L(\lambda)$.   

    Hint: please refer to problem 3 from groupwork 5 for dealing with the four light bulbs.

b. Derive an expression for the MLE of $\lambda_0$, the true value of the rate parameter. Also verify it is the global maximum of the likelihood function. Be sure to show steps.
  
3. (Simple linear regression) Suppose $X_1,X_2,\dots,X_n \overset{ind}{\sim} Norm(a_i \mu, 1)$ where the $a_i$ are \textbf{known constants} and $\mu$ is an unknown parameter. (FYI: This is an example of a linear regression model since the mean of $X$ is assumed to be a linear function of the $a_i$ )

a. Write the likelihood function $L(\mu)$ and also the log-likelihood function $\ell(\mu)$ when we observe $x_1,x_2,\dots,x_n$.

b. Derive an expression for $\widehat{\mu}_0^{mle}$, the MLE for $\mu_0$ the true value of the parameter.  (Please show your steps clearly. No need for the the second derivative test)

c. Is $\widehat{\mu}_0^{mle}$ an unbiased estimator of $\mu_0$? Show your work.

d. Derive the standard error of $\widehat{\mu}_0^{mle}$. Show your work. 

